---
title: "20-data-exploration"
output: html_notebook
---

The purpose of this notebook is for EDA.

#Load and run functionality from loading notebook
```{r source data files}
source(knitr::purl("10-load-data.Rmd"))
fs::file_delete("10-load-data.R")
```

# Violin Plotting
Let's begin creating violin plots for variables of interested identified in previous studies
```{r}
lwviolin <- ggplot(artifact_data, aes(x= particle_class, y=l_w_ratio)) + 
  geom_violin()

circleviolin <- ggplot(artifact_data, aes(x= particle_class, y=circularity)) + 
  geom_violin()

transpviolin <- ggplot(artifact_data, aes(x= particle_class, y=transparency)) + 
  geom_violin()

angulviolin <- ggplot(artifact_data, aes(x= particle_class, y=angularity)) + 
  geom_violin()

solidity_violin <- ggplot(artifact_data, aes(x= particle_class, y=solidity)) + 
  geom_violin()

grid.arrange(lwviolin,circleviolin,transpviolin ,angulviolin,solidity_violin,  nrow = 2)

```
Here we notice a significant difference between the distributions for transparency and circularly, particularly with the modes being so different. 
Both solidity and length to width ratio may be candidates but both suffer from significant outliers skewing the plots
Finally, angularity has an interesting bimodal distribution which raises questions why both experimental and site particles experience similarly peaked bimodal distributions

#Histogram Variable comparisons

Let's use the data explorer package on the experimental data

```{r}
artifact_data %>%
  filter(particle_class == "exp" ) %>%
  plot_histogram() 

artifact_data %>%
  filter(particle_class == "exp" ) %>%
  plot_density()

# Full Report Creation (optional)
# artifact_data %>%
#  filter(particle_class == "exp" ) %>%
#  create_report()

```

And compare the output to what it looks like for the site data
```{r}
artifact_data %>%
  filter(particle_class == "site" ) %>%
  plot_histogram() 

artifact_data %>%
  filter(particle_class == "site" ) %>%
  plot_density()
```
