---
title: "40-modeling"
output: html_document
---

In this notebook, we'll model the data using appropriate methods.  Here, we will add some of the generalized functions for modeling.

**Note that this notebook requires 10 to already have been run.**

```{r modeling imports}
pacman::p_load(tidymodels, usemodels)
```

# Split the data
In this step, we perform a general data split as normal.  We'll split with 75% to training and the other 25% to testing, stratified by the class.  Note that we want to abstain from looking at the performance on the test data until the _very, VERY_ end.

```{r ml data split}
set.seed(2434)
data_split <- initial_split(artifact_data, prop=3/4, strata=particle_class)

train_data <- training(data_split)
test_data <- testing(data_split)
```


# Add cross-validation splits for tuning
Here, I'll add the cross validation splits for tuning.  We'll use 5 fold cross-validation here, and we'l stratify by particle class one again.
```{r}
cv_folds <- vfold_cv(train_data, v=5, strata=particle_class)
cv_folds
```
